A Switching Planner for Combined Task and Observation Planning

### 2011-05-01 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Bunch of errors in some of the sentences in the switching section. Now seem to be fixed.

Because online DT planning is impractical for the size of problem we
are interested in, DT sessions plan in a small abstract problem
defined in terms of the trace from the proceeding sequential
session. This abstract state-space is characterised by a limited
number of propositions, chosen due to their {\em relevance} to the
action whose scheduled execution triggered the switch to the DT
session.  To allow the DT planner to judge assumptions from the trace,
we add {\em disconfirm} and {\em confirm} actions to the problem for
each of them. Those yield a relatively small reward/penalty if the
corresponding judgement is true/false. If a judgement action is
scheduled for execution, then the DT session is terminated, and a new
sequential session begins.

@@

To describe the optimisation criteria used during sequential sessions
we model $\assumptiveS{i}$ probabilistically, supposing that its
application in state $\state$ effects a transition to $\state^{T_i}$
with probability $\prob_i$, and to $\state^\bot$ with probability $1 -
\prob_i$. State $\state^\bot$ is an added sink. Now taking $\prob_i$
be the probability that the $i^{th}$ sequenced action, $\action_i$,
from a trace of state-action pairs $\langle \state_0,
\action_0,\state_1, \action_1,.., \state_N \rangle$ does not
transition to $\state^\bot$, then the optimal sequential plan has
value:

@@

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a small positive
(negative) numeric quantity which captures the utility the agent
receives for correctly (incorrectly) rejecting an assumption.  In
terms of action physics, a disconfirm action can only be executed
once, and otherwise is modelled as a self-transformation.  We only
consider {\em relevant} assumptions when constructing the abstract
model.  If \switchAction\ is the action that switched the system to a
DT session, then an assumption $\assumptiveS{i}$ is {\em relevant} if
it is necessary for the outcome of \switchAction\ to be
determined. For example, taking the switching action \switchAction\ to
be $(\pp{look}~\pp{box}~\pp{kitchen})$ from the example sequential
plan, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}))$ is not
relevant, and therefore we exclude the corresponding disconfirm action
from the abstract decision process. Given \switchAction, we also
include another once-only self-transition action
$\actions.\poss(\switchAction)$, a \emph{confirmation action} with the
reward property:

@@

In order to guarantee fast DT sessions, those plan in an abstract
process determined by the current trace and underlying belief-state.
The abstract process posed to the DT planner is constructed by first
constraining as statically false all propositions except those which
are true with probability 1, or which are the subject of {\em
  relevant} assumptions. For example, taking the above trace and
switching action {\texttt \mbox{(look~box~office)}}, the underlying
belief in Fig.~\ref{fig:beliefs}B would determine a fully constrained
belief given by Fig.~\ref{fig:beliefs}A.  Next, static constraints are
removed, one proposition at a time, until the number of states that
can be true with non-zero probability in the initial belief of the
abstract process reaches a given threshold.  In detail, for each
statically-false proposition we compute the {\em entropy} of the state
assumptions of the current trace {\em conditional} on that
proposition.  Let $X$ be a set of propositions and $2^X$ the powerset
of $X$, then taking

@@

\begin{figure}
\tiny
\begin{tabularx}{\columnwidth}{XX}
\hline
\scriptsize\vspace{0.5ex}(A) Fully constrained belief & 
\vspace{0.5ex}\hspace{-4.0ex}(C) Partially constrained belief\\
\vspace{-3.0ex}\begin{tabtt}
(\=:init (=(is-in~Robot)kitchen) \+ \\
       (\=.6(=(is-in~box)kitchen))) \\
\end{tabtt} & 
\vspace{-3.0ex}\begin{tabtt}
\hspace{-5.0ex}(\=:init (=(is-in~Robot)kitchen) \+ \\
       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
        (\=.9(=(is-in~milk)~kitchen))\+\\
         .1(=(is-in~milk)office))\-\-  \\
       .4(and(=(is-in~box)office) \+\\
        (.1(=(is-in~milk)kitchen))\+\\
        .9(=(is-in~milk)office)))
\end{tabtt}\\
\vspace{-16.5ex}(B) Underlying DTPDDL belief & \\ 
\vspace{-17.0ex}\begin{tabtt}
(\=:init (=(is-in~Robot)office) \+ \\
       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
        (\=.9(=(is-in~milk)kitchen))\+\\
         .1(=(is-in~milk)office))\-\-  \\
       .4(and(=(is-in~box)office) \+\\
        (.1(=(is-in~milk)kitchen))\+\\
        .9(=(is-in~milk)office))) \-\-\-\\
       (\=.6(=(is-in~cup)office) \+ \\
      .4(=(is-in~cup)kitchen))) \\
\end{tabtt} & \\
\hline
\end{tabularx}
\caption{Simplified examples of abstract belief-states from
       DT sessions. \label{fig:beliefs}}
\normalsize
\end{figure}

>>>>

%% due to their {\em relevance} to the
%% action whose scheduled execution triggered the switch to the DT
%% session.

Because online DT planning is impractical for the size of problem we
are interested in, DT sessions plan in a small abstract problem
defined in terms of the trace from the proceeding sequential session.
This abstract state-space is characterised by a limited number of
propositions, chosen because they relate evidence about assumptions
in the trace.  To allow the DT planner to judge {\em relevant}
assumptions from the trace, we add {\em disconfirm} and {\em confirm}
actions to the problem for each of them. Those yield a relatively
small reward/penalty if the corresponding judgement is true/false. If
a judgement action is scheduled for execution, then the DT session is
terminated, and a new sequential session begins.

@@

To describe the optimisation criteria used during sequential sessions
we model $\assumptiveS{i}$ probabilistically, supposing that its
application in state $\state$ effects a transition to $\state^{T_i}$
with probability $\prob_i$, and to $\state^\bot$ with probability $1 -
\prob_i$. State $\state^\bot$ is an added sink. Taking $\prob_i$ to be
the probability that the $i^{th}$ sequenced action, $\action_i$, from
a trace of state-action pairs $\langle \state_0, \action_0,\state_1,
\action_1,.., \state_N \rangle$ does not transition to $\state^\bot$,
then the optimal sequential plan has value:

@@

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a small positive
(negative) numeric quantity which captures the utility the agent
receives for correctly (incorrectly) rejecting an assumption.  In
terms of action physics, a disconfirm action can only be executed
once, and otherwise is modelled as a self-transformation.  We only
consider {\em relevant} assumptions when constructing the abstract
model.  If \switchAction\ is the action that switched the system to a
DT session, then an assumption $\assumptiveS{i}$ is {\em relevant} if
it is necessary for the outcome of \switchAction\ to be
determined. For example, taking the switching action \switchAction\ to
be $(\pp{look}~\pp{box}~\pp{kitchen})$ from our earlier sequential
plan example, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}))$ is not
relevant, and therefore we exclude the corresponding disconfirm action
from the abstract decision process. Given \switchAction, we also
include another once-only self-transition action
$\actions.\poss(\switchAction)$, a \emph{confirmation action} with the
reward property:

@@

In order to guarantee fast DT sessions, those plan in an abstract
process determined by the current trace and underlying belief-state.
The abstract process posed to the DT planner is constructed by first
constraining as statically false all propositions except those which
are true with probability 1, or which are the subject of {\em
  relevant} assumptions. For example, taking the above trace and
switching action {\texttt \mbox{(look~box~office)}}, the underlying
belief in Fig.~\ref{fig:beliefs}B would determine a fully constrained
belief given by Fig.~\ref{fig:beliefs}A.  Next, static constraints are
removed, one proposition at a time, until the number of states that
can be true with non-zero probability in the initial belief of the
abstract process reaches a given threshold.  In detail, for each
statically-false proposition we compute the {\em entropy} of the
relevant assumptions of the current trace {\em conditional} on that
proposition.  Let $X$ be a set of propositions and $2^X$ the powerset
of $X$, then taking

@@

\begin{figure}
\tiny
\begin{tabularx}{\columnwidth}{XX}
\hline
\scriptsize\vspace{0.5ex}(A) Fully constrained belief & 
\vspace{0.5ex}\hspace{-4.0ex}(C) Partially constrained belief\\
\vspace{-3.0ex}\begin{tabtt}
(\=:init (=(is-in~Robot)kitchen) \+ \\
       (\=.6(=(is-in~box)kitchen))) \\
\end{tabtt} & 
\vspace{-3.0ex}\begin{tabtt}
\hspace{-5.0ex}(\=:init (=(is-in~Robot)kitchen) \+ \\
       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
        (\=.9(=(is-in~milk)~kitchen))\+\\
         .1(=(is-in~milk)office))\-\-  \\
       .4(and(=(is-in~box)office) \+\\
        (.1(=(is-in~milk)kitchen))\+\\
        .9(=(is-in~milk)office)))
\end{tabtt}\\
\vspace{-16.5ex}(B) Underlying DTPDDL belief & \\ 
\vspace{-17.0ex}\begin{tabtt}
(\=:init (=(is-in~Robot)kitchen) \+ \\
       (\=.6(a\=nd(=(is-in~box)kitchen) \+\+ \\
        (\=.9(=(is-in~milk)kitchen))\+\\
         .1(=(is-in~milk)office))\-\-  \\
       .4(and(=(is-in~box)office) \+\\
        (.1(=(is-in~milk)kitchen))\+\\
        .9(=(is-in~milk)office))) \-\-\-\\
       (\=.6(=(is-in~cup)office) \+ \\
      .4(=(is-in~cup)kitchen))) \\
\end{tabtt} & \\
\hline
\end{tabularx}
\caption{Simplified examples of belief-states from
       DT sessions. \label{fig:beliefs}}
\normalsize
\end{figure}
